### 2024-12-24 Article/main.tex
committed by yang: add LaTeX files of article and data for calculation

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% --------------------------------------------------------
% Rho
% LaTeX Template
% Version 2.1.1 (01/09/2024)
%
% Authors: 
% Guillermo Jimenez (memo.notess1@gmail.com)
% Eduardo Gracidas (eduardo.gracidas29@gmail.com)
% 
% License:
% Creative Commons CC BY 4.0
% --------------------------------------------------------
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[9pt,a4paper,twoside]{rho-class/rho}
\usepackage[english]{babel}
% \usepackage{ctex}

%% Spanish babel recomendation
% \usepackage[spanish,es-nodecimaldot,es-noindentfirst]{babel}

\setbool{rho-abstract}{true} % Set false to hide the abstract
\setbool{corres-info}{true} % Set false to hide the corresponding author section

%----------------------------------------------------------
% TITLE
%----------------------------------------------------------

\journalname{Central South University}
\title{Geoelectric Field and Electrical Exploration Assignment: Time Series Signal Processing in Magnetotelluric Method}

%----------------------------------------------------------
% AUTHORS AND AFFILIATIONS
%----------------------------------------------------------

\author[1]{Yaokun Yang}
\author[1]{Senyu Zeng}
\author[1]{Binyan Wei}
\author[1]{Qiuying Zhao}
\author[1]{Haoyu Zhang}

%----------------------------------------------------------

\affil[1]{Shcool of Geoscience and Info-physics, Central South University, Changsha 410083, China.}

%----------------------------------------------------------
% DATES
%----------------------------------------------------------

\dates{This manuscript was compile on December 24, 2024}

%----------------------------------------------------------
% FOOTER INFORMATION
%----------------------------------------------------------

\leadauthor{Y. Yang et al.}
\footinfo{Advisor: Prof. Jiabin Yan}
\smalltitle{\LaTeX\ Template}
\institution{Central South University}
\theday{December 24, 2024} %\today

%----------------------------------------------------------
% ARTICLE INFORMATION
%----------------------------------------------------------

\corres{Yaokun Yang, Department of Applied Geophysics.}
\email{yangyaokun@csu.edu.cn.}
\doi{\url{https://www.doi.org/fakedoi/585858558}}

\received{December 24, 2024}
\revised{December 24, 2024}
\accepted{December 25, 2024}
\published{December 26, 2024}

\license{Adhere to the MIT license.}

%----------------------------------------------------------
% ABSTRACT
%----------------------------------------------------------

\begin{abstract}

\end{abstract}

%----------------------------------------------------------

\keywords{magnetotelluric method, signal processing, least square method}

%----------------------------------------------------------

\begin{document}

    \maketitle
    \thispagestyle{firststyle}
    % \tableofcontents
    \linenumbers

%----------------------------------------------------------

\section{Introduction}

\section{Methodology}

\section{Algorithm and implementation}

\section{Result}

\section{Conclusions}

    % \begin{rhoenv}[frametitle=Environment with custom title]
    %     Hello! I am an example of the \textit{rhoenv} included in rhoenvs \LaTeX\ package. Here you can include relevant information or notes about your work. You can modify my title directly in the code.
    % \end{rhoenv}

    % \begin{figure}[H]
    %     \centering
    %     \includegraphics[width=0.71\columnwidth]{Example.pdf}
    %     \caption{Example figure obtained from PGFPlots \cite{PFGPlots}.}
    %     \label{fig:figure}
    % \end{figure}

    % \begin{table*}[pb]
    %     \RaggedRight
    %     \caption{Table example that covers the width of the page.}
    %     \label{tab:table}
    %         \begin{tabular}{lllp{12.2cm}}
    %             \toprule
    %             \textbf{Day} & \textbf{Min Temp} & \textbf{Max Temp} & \textbf{Summary} \\ 
    %             \midrule
    %             Monday & 11\textdegree C & 22\textdegree C & A clear day with lots of sunshine.  Strong breeze will bring down the temperatures. \\
    %             Tuesday & 9\textdegree C & 19\textdegree C & Cloudy with rain, across many northern regions. \\
    %             Wednesday & 10\textdegree C & 21\textdegree C & Rain will still linger for the morning. 
    %             Conditions will improve by early afternoon and continue 
    %             throughout the evening.\\
    %             \bottomrule
    %         \end{tabular}

    %     \tabletext{Note: Obtained from \LaTeX\ tables \cite{projects-2023}.}

    % \end{table*}

    % \nolinenumbers
    % \lstinputlisting[caption=Example of matlab code., label={lst:listing-Mat}, language=Matlab]{example.m}
    % \linenumbers

    % \nolinenumbers
    % \begin{lstlisting}[language=TeX, caption=Alternative unnumbered section.]
    % \titleformat{name=\section,numberless}[block]
    %     {\color{rhocolor}\sffamily\large\bfseries}
    %     {}
    %     {0em}
    %     {#1}
    %     []
    % \end{lstlisting}
    % \linenumbers

%----------------------------------------------------------

\printbibliography

%----------------------------------------------------------

\end{document}